Gender and Mental Depression: Analyzing Census Data Through Statistical Methods

### 2024-10-22 paper/paper.qmd
Update paper.qmd
Updated #Introduction to include an overview of the paper's scope, variables used for training the model, and a short paragraph on why  models like this matter.

Overview paragraph

@@

Why it matters paragraph

>>>>

Mental health issues, particularly depression, have emerged as a significant public health concern globally. Among other factors influencing mental health, [gender and education] stand out as critical determinants. Recent studies have also indicated that educational attainment may play a protective role against depression, suggesting that individuals with higher education levels experience lower rates of mental health issues. Conversely, gender disparities in depression prevalence remain pronounced, with research consistently showing that women report higher rates of depression than their male counterparts, even when controlling for education levels.
Therefore our key questions underpinning this research include: does higher education correlate with lower rates of depression? Are women disproportionately affected by depression compared to men at the same educational level? Can we develop a model that predicts the likelihood of a [KSUM score of 13 or higher] given certain characteristics of an individual [please list the characteristics we used here, and what we called them in your data or graphs]? To address these questions, we analyze national census data through various statistical methods, focusing on the intersection of gender, education, income, and mental health outcomes [please cite the IPUMs dataset here, with a footnote on how you selected your filters].

This study employs a statistical approach in R  to examine the relationships between gender, educational attainment, income and reported cases of mental depression. By exploring these dynamics, we hope to contribute to a deeper understanding of how education and gender intersect to influence mental health, providing valuable insights for policymakers and mental health practitioners.

The remainder of this paper is structured as follows: Section 2 presents the data and methodology, detailing the statistical techniques used for analysis. Section 3 discusses the modeling strategy and expected outcomes, followed by a presentation of results in Section 4. Finally, Section 5 engages in a critical discussion of our findings, addressing their implications for future research and policy.

@@

Why it matters paragraph
Predictive models - even simplified ones like the one we present below - are helpful for understanding complex issues like mental health. They allow researchers and policymakers spot trends and make smarter decisions. By looking at how factors like gender and education connect to depression, these models show patterns which lets us identify at-risk groups early, and lets us provide tailored support and resources to those who need it most. Plus, predictive models help direct future research by pointing out where more exploration is needed, ensuring we keep learning and improving our approach to mental health. In a time when mental well-being is more important than ever, being able to predict and tackle what affects it is key to building healthier communities.
